

```

### Approximation 1:


```{r}
library(ggplot2)

# Give a value for the median
m<-0.5

# define exponential rate/intensity parameter
lambda<-log(2)/m

# set percentile to model (as a proportion)
p<-0.5

# find median of exponential
median<-log(2)/lambda

# find median on log scale
medianz<-log(median)

# generate a sample from the exponential distribution
y<-rexp(n=1000,rate=lambda)

# same sample on the log scale
z<-log(y)

# True density and survival function
fy<-lambda*exp(-lambda*y)
Sy<-exp(-lambda*y)
# Calculate approximation 1 to the density
# Checked these derivatives on 11/11/2018

# define Tau=logit(F(quantile))
Tau<-log(p/(1-p))

# find value of quantile
q<--log(1-p)/lambda

# first order Taylor approximation to LF at the given quantile
LF1qy<-1/(1-exp(-q*lambda))*lambda
# second order Taylor approximation to LF at the given quantile
LF2qy<--1/(1-exp(-q*lambda))^2*lambda^2*exp(-q*lambda)
# third order Taylor approximation to LF at the given quantile
LF3qy<-2/(1-exp(-q*lambda))^3*exp(-q*lambda)^2*lambda^3+1/(1-exp(-q*lambda))^2*lambda^3*exp(-q*lambda)

# approximation to the density using the first order Taylor approximation for LF
f1qy<-1/(1+exp(-(Tau+(y-q)*LF1qy)))*1/(1+exp(Tau+(y-q)*LF1qy))*LF1qy
# approximation to the density using the second order Taylor approximation for LF
f2qy<-1/(1+exp(-(Tau+(y-q)*LF1qy+LF2qy*(y-q)^2/2)))*1/(1+exp(Tau+(y-q)*LF1qy+LF2qy*(y-q)^2/2))*(LF1qy+LF2qy*(y-q))
# approximation to the density using the third order Taylor approximation for LF
f3qy<-1/(1+exp(-(Tau+(y-q)*LF1qy+LF2qy*(y-q)^2/2+LF3qy*(y-q)^3/6)))*1/(1+exp(Tau+(y-q)*LF1qy+LF2qy*(y-q)^2/2+LF3qy*(y-q)^3/6))*(LF1qy+LF2qy*(y-q)+LF3qy*(y-q)^2/2)

# approximation to the survival function using the first order Taylor approximation for LF
S1qy<-1/(1+exp(Tau+(y-q)*LF1qy))
# approximation to the survival function using the second order Taylor approximation for LF
S2qy<-1/(1+exp(Tau+(y-q)*LF1qy+LF2qy*(y-q)^2/2))
# approximation to the survival function using the third order Taylor approximation for LF
S3qy<-1/(1+exp(Tau+(y-q)*LF1qy+LF2qy*(y-q)^2/2+LF3qy*(y-q)^3/6))

# Density Plots of different orders of approximation 1 
figure1.a<-ggplot()+
  ggtitle('Figure 1.a. Approximation 1: 1st-, 2nd-, and 3rd-order Density Approximations')+
  labs(y= "Density of Y", x = "y")+
  scale_colour_manual("", 
  values = c("True Density"="black","First-Order"="blue","Second-Order"="green","Third-Order"="red"))+
  geom_line(aes(y,fy, color="True Density"))+
  geom_line(aes(y,f1qy, color='First-Order'))+
  geom_line(aes(y,f2qy, color='Second-Order'))+
  geom_line(aes(y,f3qy, color='Third-Order'))

figure1.a

# Survival function Plots of different orders of approximation 1 
figure1.b<-ggplot()+
  ggtitle('Figure 1.b. Approximation 1: 1st-, 2nd-, and 3rd-order Survivor Function Approximations')+
  labs(y= "Survivor Function of Y", x = "y")+
  scale_colour_manual("", 
                      values = c("True Survivor Function"="black","First-Order"="blue","Second-Order"="green","Third-Order"="red"))+
  geom_line(aes(y,Sy, color="True Survivor Function"))+
  geom_line(aes(y,S1qy, color='First-Order'))+
  geom_line(aes(y,S2qy, color='Second-Order'))+
  geom_line(aes(y,S3qy, color='Third-Order'))

figure1.b


# Calculate densities for approximation 2
LF1y<--2*log(2)/y
LF2y<-2*(log(2)/(y^2))*(2-y*log(2))
LF3y<-2*log(2)/y^3*(2-y*log(2))*(y*log(2)-3)

hcy<--LF1y
hc1y<-2*(log(2)/y)^2
hc2y<-2*(log(2))^2/y^3*(y*log(2)-3)

f1y<-1/(1+exp(-(median-y)*LF1y))*1/(1+exp((median-y)*LF1y))*hcy
f2y<-1/(1+exp(-(median-y)*(LF1y+LF2y*(median-y)/2)))*1/(1+exp((median-y)*(LF1y+LF2y*(median-y)/2)))*(hcy+hc1y*(median-y))
f3y<-1/(1+exp(-(median-y)*(LF1y+LF2y*(median-y)/2+LF3y*(median-y)^2/6)))*1/(1+exp((median-y)*(LF1y+LF2y*(median-y)/2+LF3y*(median-y)^2/6)))*(hcy+hc1y*(median-y)+hc2y*(median-y)^2/2)

S1y<-1/(1+exp((median-y)*LF1y))
S2y<-1/(1+exp((median-y)*(LF1y+LF2y*(median-y)/2)))
S3y<-1/(1+exp((median-y)*(LF1y+LF2y*(median-y)/2+LF3y*(median-y)^2/6)))

# Plots of different orders of approximation 2 
figure2.a<-ggplot()+
  ggtitle('Figure 2.a. Approximation 2: 1st-, 2nd-, and 3rd-order Density Approximations')+
  labs(y= "Density of Y", x = "y")+
  scale_colour_manual("", 
                      values = c("True Density"="black","First-Order"="blue","Second-Order"="green","Third-Order"="red"))+
  geom_line(aes(y,fy, color="True Density"))+geom_line(aes(y,f1y, color='First-Order'))+
  geom_line(aes(y,f2y, color='Second-Order'))+
  geom_line(aes(y,f3y, color='Third-Order'))

figure2.a

# Survival function Plots of different orders of approximation 2 
figure2.b<-ggplot()+
  ggtitle('Figure 2.b. Approximation 2: 1st-, 2nd-, and 3rd-order Survivor Function Approximations')+
  labs(y= "Survivor Function of Y", x = "y")+
  scale_colour_manual("", 
                      values = c("True Survivor Function"="black","First-Order"="blue","Second-Order"="green","Third-Order"="red"))+
  geom_line(aes(y,Sy, color="True Survivor Function"))+
  geom_line(aes(y,S1y, color='First-Order'))+
  geom_line(aes(y,S2y, color='Second-Order'))+
  geom_line(aes(y,S3y, color='Third-Order'))

figure2.b

# Calculate densities for approximation 2 of transformed data (intermediate step in approximation 4)
LF1z=-2*log(2)
LF2z=-2*log(2)+2*(log(2)^2)
LF3z=-2*log(2)+6*(log(2)^2)-6*(log(2)^3)

hcz<--LF1z
hc1z<--LF2z
hc2z<--LF3z

fz<-fy*y
f1z<-1/(1+exp(-(medianz-z)*LF1z))*1/(1+exp((medianz-z)*LF1z))*hcz
f2z<-1/(1+exp(-(medianz-z)*(LF1z+LF2z*(medianz-z)/2)))*1/(1+exp((medianz-z)*(LF1z+LF2z*(medianz-z)/2)))*(hcz+hc1z*(medianz-z))
f3z<-1/(1+exp(-(medianz-z)*(LF1z+LF2z*(medianz-z)/2+LF3z*(medianz-z)^2/6)))*1/(1+exp((medianz-z)*(LF1z+LF2z*(medianz-z)/2+LF3z*(medianz-z)^2/6)))*(hcz+hc1z*(medianz-z)+hc2z*(medianz-z)^2/2)

S1z<-1/(1+exp((medianz-z)*LF1z))
S2z<-1/(1+exp((medianz-z)*(LF1z+LF2z*(medianz-z)/2)))
S3z<-1/(1+exp((medianz-z)*(LF1z+LF2z*(medianz-z)/2+LF3z*(medianz-z)^2/6)))

# Plots of different orders of approximation 2 of transformed data (intermediate step in approximation 4)
figure5.a<-ggplot()+
  ggtitle('Figure 5.a. Approximation 4, Step 1: 1st-, 2nd-, and 3rd-order \n Density Approximations for Transformed Data')+
  labs(y= "Density of log(Y)", x = "log(y)")+
  scale_colour_manual("", 
                      values = c("True Density"="black","First-Order"="blue","Second-Order"="green","Third-Order"="red"))+
  geom_line(aes(z,fz, color="True Density"))+
  geom_line(aes(z,f1z, color='First-Order'))+
  geom_line(aes(z,f2z, color='Second-Order'))+
  geom_line(aes(z,f3z, color='Third-Order'))

figure5.a

# Survival function Plots of different orders of approximation 2 of transformed data 
figure5.b<-ggplot()+
  ggtitle('Figure 5.b. Approximation 4, Step 1: 1st-, 2nd-, and 3rd-order Survivor Function Approximations')+
  labs(y= "Survivor Function of Y", x = "y")+
  scale_colour_manual("", 
                      values = c("True Survivor Function"="black","First-Order"="blue","Second-Order"="green","Third-Order"="red"))+
  geom_line(aes(z,Sy, color="True Survivor Function"))+
  geom_line(aes(z,S1z, color='First-Order'))+
  geom_line(aes(z,S2z, color='Second-Order'))+
  geom_line(aes(z,S3z, color='Third-Order'))

figure5.b
```

### Approximation 4, Step 2: Back-Transform Results from Step 1

Using the transformation theorem, we obtain the following densities for the original expenentially-distributed random variable Y:

```{r}
# Calculate densities for back-transformed data, giving approximation 4
g1y<-f1z/y
g2y<-f2z/y
g3y<-f3z/y

# Plots of different orders of approximation 4
figure6.a<-ggplot()+
  ggtitle('Figure 6. Approximation 4: 1st-, 2nd-, and 3rd-order Density Approximations')+
  labs(y= "Density of Y", x = "y")+
  scale_colour_manual("", 
                      values = c("True Density"="black","First-Order"="blue","Second-Order"="green","Third-Order"="red"))+
  geom_line(aes(y,fy, color="True Density"))+
  geom_line(aes(y,g1y, color='First-Order'))+
  geom_line(aes(y,g2y, color='Second-Order'))+
  geom_line(aes(y,g3y, color='Third-Order'))

figure6.a

# Survival function Plots of different orders of approximation 1 
figure6.b<-ggplot()+
  ggtitle('Figure 6.b. Approximation 4: 1st-, 2nd-, and 3rd-order Survivor Function Approximations')+
  labs(y= "Survivor Function of Y", x = "y")+
  scale_colour_manual("", 
                      values = c("True Survivor Function"="black","First-Order"="blue","Second-Order"="green","Third-Order"="red"))+
  geom_line(aes(y,Sy, color="True Survivor Function"))+
  geom_line(aes(y,S1z, color='First-Order'))+
  geom_line(aes(y,S2z, color='Second-Order'))+
  geom_line(aes(y,S3z, color='Third-Order'))

figure6.b
```

## Comparison of First-Order Approximations

The following plots overlay the first-order approximations for approximations 1 through 4 for comparison.

```{r}
# Plots of different 1st-order approximations
figure7.a<-ggplot()+
  ggtitle('Figure 7.a. 1st-Order Density Approximations')+
  labs(y= "Density of Y", x = "y")+
  scale_colour_manual("", 
                      values = c("True Density"="black","Approximation 1"="blue","Approximation 2"="green","Approximation 3"="red","Approximation 4"="purple"))+
  geom_line(aes(y,fy, color="True Density"))+
  geom_line(aes(y,f1my, color='Approximation 1'))+
  geom_line(aes(y,f1y, color='Approximation 2'))+
  geom_line(aes(y,g1my, color='Approximation 3'))+
  geom_line(aes(y,g1y, color='Approximation 4'))

figure7.a

# Plots of different 1st-order approximations
figure7.b<-ggplot()+
  ggtitle('Figure 7.b. 1st-Order Survivor Function Approximations')+
  labs(y= "Survivor Function of Y", x = "y")+
  scale_colour_manual("", 
                      values = c("True Survivor Function"="black","Approximation 1"="blue","Approximation 2"="green","Approximation 3"="red","Approximation 4"="purple"))+
  geom_line(aes(y,Sy, color="True Survivor Function"))+
  geom_line(aes(y,S1my, color='Approximation 1'))+
  geom_line(aes(y,S1y, color='Approximation 2'))+
  geom_line(aes(y,S1mz, color='Approximation 3'))+
  geom_line(aes(y,S1z, color='Approximation 4'))

figure7.b

Median<-matrix(nrow=1000,ncol=1)

for (i in 1:1000){
  Median[i,1]<-i/500
}

yTrans<-matrix(nrow=1000,ncol=1000)
for (i in 1:1000){
  yTrans[i,]<-t(y)
}

trueLL<-matrix(nrow=1000,ncol=1)
LLm1<-matrix(nrow=1000,ncol=1)
LLy1<-matrix(nrow=1000,ncol=1)
LLLm1<-matrix(nrow=1000,ncol=1)
trueMedian<-matrix(nrow=1000,ncol=1)
for (i in 1:1000){
  trueLL[i,1]<-0
  LLm1[i,1]<-0
  LLy1[i,1]<-0
  LLLm1[i,1]<-0  
  trueMedian[i,1]<-median
}

for (i in 1:1000){
  trueLL<-trueLL+log(log(2)/Median*exp(-yTrans[,i]*log(2)/Median))
  LLm1<-LLm1+log(1/(1+exp(-(yTrans[,i]-Median)*2*log(2)/Median))*1/(1+exp((yTrans[,i]-Median)*2*log(2)/Median))*2*log(2)/Median)
  LLy1<-LLy1+log(1/(1+exp(-(Median-yTrans[,i])*(-2)*log(2)/yTrans[,i]))*1/(1+exp((Median-yTrans[,i])*(-2)*log(2)/yTrans[,i]))*(2)*log(2)/yTrans[,i])
  
  LLLm1<-LLLm1+
    log(
      1/(1+exp(
        -(log(yTrans[,i])-log(Median))*
          (LF1z)
      ))*
        1/(1+exp(
          (log(yTrans[,i])-log(Median))*
            (LF1z)
        ))*
        (hcz)
      /yTrans[,i]
    )
}


# Plots of different 1st-order approximations to the likelihood
figure8<-ggplot()+ylim(-700,900)+geom_line(aes(trueMedian,trueLL, color="True Median"))+
  ggtitle('Figure 8. 1st-Order Log-Likelihood Approximations')+
  labs(y= "Log-Likelihood of the Median", x = "Median")+
  scale_colour_manual("", 
                      values = c("True Median"="gray","True Log-Likelihood"="black","Approximation 1"="blue"))+
  geom_line(aes(Median,trueLL, color="True Log-Likelihood"))+
  geom_line(aes(Median,LLm1, color='Approximation 1'))

figure8
